

```{r dataprep, message = F, warning = F, cache = T}
# Datasets
load(here("Output/anz_euro_pheno_prs.RData"))
load(here("Output/clear_lasso_euro_pheno_prs.RData"))
load(here("Output/control_anz_euro_pheno_prs.RData"))
load(here("Output/control_eastpoly_pheno_prs.RData"))
load(here("Output/control_westpoly_pheno_prs.RData"))
load(here("Output/crystal_euro_pheno_prs.RData"))
load(here("Output/eastpoly_pheno_prs.RData"))
load(here("Output/eurogout_euro_pheno_prs.RData"))
load(here("Output/lasso_other_euro_pheno_prs.RData"))
load(here("Output/light_euro_pheno_prs.RData"))
load(here("Output/westpoly_pheno_prs.RData"))
load(here("Output/ukbb_euro_pheno_prs.RData"))
load(here("Output/ukbb_poly_pheno_prs.RData"))

load(here("Output/UKBB_Gene_OR.RData"))
load(here("Output/Poly_Gene_OR.RData"))


# European cohort prep
anz_gout <- anz_euro_pheno_prs %>% 
  mutate(PRS_std = (PRS - mean(PRS)) / sd(PRS),
         FLAREHIGH = if_else(NUMATK > 1, TRUE, FALSE),
         COHORT2 = "Aus/NZ") %>% 
  select(-Geno.SampleID)

anz_control <- control_anz_euro_pheno_prs %>%
  mutate(PRS_std = (PRS - mean(PRS)) / sd(PRS),
         GOUT = FALSE,
         SEX = factor(Pheno.Gender, levels = c("Female", "Male")),
         SURICACID = Pheno.CollectionUrate / 0.05949) %>% 
  rename(IID = Pheno.SampleID,
         AGECOL = Pheno.Age,
         COHORT = Pheno.Study,
         PC1 = Geno.PCVector1,
         PC2 = Geno.PCVector2,
         PC3 = Geno.PCVector3,
         PC4 = Geno.PCVector4,
         PC5 = Geno.PCVector5,
         PC6 = Geno.PCVector6,
         PC7 = Geno.PCVector7,
         PC8 = Geno.PCVector8,
         PC9 = Geno.PCVector9,
         PC10 = Geno.PCVector10) %>% 
  select(IID, COHORT, SEX, AGECOL, GOUT, SURICACID, PC1:PC10, PRS:rs738408, PRS_std)

tmp <- anz_gout %>% 
  select(IID, COHORT, SEX, AGECOL, GOUT, SURICACID, PC1:PC10, PRS:rs738408, PRS_std)

anz_full <- rbind(tmp, anz_control) %>% 
  mutate(PRS_std = (PRS - mean(PRS)) / sd(PRS))

eurogout <- eurogout_euro_pheno_prs %>% 
  mutate(PRS_std = (PRS - mean(PRS)) / sd(PRS),
         FLAREHIGH = if_else(NUMATK > 1, TRUE, FALSE),
         COHORT2 = "EuroGout") %>% 
  select(-Geno.SampleID)


ardea_clear_lasso <- clear_lasso_euro_pheno_prs %>% 
  mutate(PRS_std = (PRS - mean(PRS)) / sd(PRS),
         FLAREHIGH = if_else(NUMATK > 1, TRUE, FALSE),
         COHORT2 = "Ardea - CLEAR/LASSO (ULT)") %>% 
  select(-Geno.SampleID)


ardea_lasso_other <- lasso_other_euro_pheno_prs %>% 
  mutate(PRS_std = (PRS - mean(PRS)) / sd(PRS),
         FLAREHIGH = if_else(NUMATK > 1, TRUE, FALSE),
         COHORT2 = "Ardea - LASSO (other)") %>% 
  select(-Geno.SampleID)


ardea_crystal <- crystal_euro_pheno_prs %>% 
  mutate(PRS_std = (PRS - mean(PRS)) / sd(PRS),
         FLAREHIGH = if_else(NUMATK > 1, TRUE, FALSE),
         COHORT2 = "Ardea - CRYSTAL") %>% 
  select(-Geno.SampleID)


ardea_light <- light_euro_pheno_prs %>% 
  mutate(PRS_std = (PRS - mean(PRS)) / sd(PRS),
         FLAREHIGH = if_else(NUMATK > 1, TRUE, FALSE),
         COHORT2 = "Ardea - LIGHT") %>% 
  select(-Geno.SampleID)
  

allgout <- rbind(anz_gout, eurogout, ardea_clear_lasso, ardea_crystal, ardea_lasso_other, ardea_light) %>% 
  mutate(PRS_std = (PRS - mean(PRS)) / sd(PRS))

rm(anz_euro_pheno_prs, clear_lasso_euro_pheno_prs, control_anz_euro_pheno_prs, crystal_euro_pheno_prs, eurogout_euro_pheno_prs, lasso_other_euro_pheno_prs, light_euro_pheno_prs, tmp)


# Polynesian cohort prep
eastpoly_gout <- eastpoly_pheno_prs %>% 
  mutate(PRS_std = (PRS - mean(PRS)) / sd(PRS),
         FLAREHIGH = if_else(NUMATK > 1, TRUE, FALSE)) %>% 
  select(-Geno.SampleID)

eastpoly_gout_tophi <- eastpoly_gout %>% 
  filter(TOPHIGOUT == TRUE)

eastpoly_gout_flare <- eastpoly_gout %>% 
  filter(FLAREHIGH == TRUE)

eastpoly_gout_onset <- eastpoly_gout %>% 
  filter(AGE1ATK < 40)

eastpoly_control <- control_eastpoly_pheno_prs %>%
  mutate(PRS_std = (PRS - mean(PRS)) / sd(PRS),
         GOUT = FALSE,
         SEX = factor(Pheno.Gender, levels = c("Female", "Male")),
         SURICACID = Pheno.CollectionUrate / 0.05949) %>% 
  rename(IID = Pheno.SampleID,
         AGECOL = Pheno.Age,
         COHORT = Pheno.Study,
         PC1 = Geno.PCVector1,
         PC2 = Geno.PCVector2,
         PC3 = Geno.PCVector3,
         PC4 = Geno.PCVector4,
         PC5 = Geno.PCVector5,
         PC6 = Geno.PCVector6,
         PC7 = Geno.PCVector7,
         PC8 = Geno.PCVector8,
         PC9 = Geno.PCVector9,
         PC10 = Geno.PCVector10,
         PC1_Oc = Geno.PCVector1_Oc,
         PC2_Oc = Geno.PCVector2_Oc,
         PC3_Oc = Geno.PCVector3_Oc,
         PC4_Oc = Geno.PCVector4_Oc,
         PC5_Oc = Geno.PCVector5_Oc,
         PC6_Oc = Geno.PCVector6_Oc,
         PC7_Oc = Geno.PCVector7_Oc,
         PC8_Oc = Geno.PCVector8_Oc,
         PC9_Oc = Geno.PCVector9_Oc,
         PC10_Oc = Geno.PCVector10_Oc) %>% 
  select(IID, COHORT, SEX, AGECOL, GOUT, SURICACID, PC1:PC10, Geno.PolynesianGroup, PC1_Oc:PC10_Oc, PRS:PRS_std)

tmp <- eastpoly_gout %>% 
  select(IID, COHORT, SEX, AGECOL, GOUT, SURICACID, PC1:PC10, Geno.PolynesianGroup, PC1_Oc:PC10_Oc, PRS:PRS_std)

eastpoly_full <- rbind(tmp, eastpoly_control) %>% 
  mutate(PRS_std = (PRS - mean(PRS)) / sd(PRS))

westpoly_gout <- westpoly_pheno_prs %>% 
  mutate(PRS_std = (PRS - mean(PRS)) / sd(PRS),
         FLAREHIGH = if_else(NUMATK > 1, TRUE, FALSE)) %>% 
  select(-Geno.SampleID)

westpoly_control <- control_westpoly_pheno_prs %>%
  mutate(PRS_std = (PRS - mean(PRS)) / sd(PRS),
         GOUT = FALSE,
         SEX = factor(Pheno.Gender, levels = c("Female", "Male")),
         SURICACID = Pheno.CollectionUrate / 0.05949) %>% 
  rename(IID = Pheno.SampleID,
         AGECOL = Pheno.Age,
         COHORT = Pheno.Study,
         PC1 = Geno.PCVector1,
         PC2 = Geno.PCVector2,
         PC3 = Geno.PCVector3,
         PC4 = Geno.PCVector4,
         PC5 = Geno.PCVector5,
         PC6 = Geno.PCVector6,
         PC7 = Geno.PCVector7,
         PC8 = Geno.PCVector8,
         PC9 = Geno.PCVector9,
         PC10 = Geno.PCVector10,
         PC1_Oc = Geno.PCVector1_Oc,
         PC2_Oc = Geno.PCVector2_Oc,
         PC3_Oc = Geno.PCVector3_Oc,
         PC4_Oc = Geno.PCVector4_Oc,
         PC5_Oc = Geno.PCVector5_Oc,
         PC6_Oc = Geno.PCVector6_Oc,
         PC7_Oc = Geno.PCVector7_Oc,
         PC8_Oc = Geno.PCVector8_Oc,
         PC9_Oc = Geno.PCVector9_Oc,
         PC10_Oc = Geno.PCVector10_Oc) %>% 
  select(IID, COHORT, SEX, AGECOL, GOUT, SURICACID, PC1:PC10, Geno.PolynesianGroup, PC1_Oc:PC10_Oc, PRS:PRS_std)

tmp <- westpoly_gout %>% 
  select(IID, COHORT, SEX, AGECOL, GOUT, SURICACID, PC1:PC10, Geno.PolynesianGroup, PC1_Oc:PC10_Oc, PRS:PRS_std)

westpoly_full <- rbind(tmp, westpoly_control) %>% 
  mutate(PRS_std = (PRS - mean(PRS)) / sd(PRS))

rm(control_eastpoly_pheno_prs, control_westpoly_pheno_prs, eastpoly_pheno_prs, westpoly_pheno_prs, tmp)


# UK Biobank prep
ukbb_gout <- ukbb_euro_pheno_prs %>% 
  filter(GOUT == TRUE,
         !is.na(PRS)) %>% 
  mutate(PRS_std = (PRS - mean(PRS)) / sd(PRS),
         FLAREHIGH = NA,
         NUMATK = NA,
         TOPHIGOUT = NA,
         AGE1ATK = NA,
         DURATION = NA,
         PC1 = NA,
         PC2 = NA,
         PC3 = NA,
         PC4 = NA,
         PC5 = NA,
         PC6 = NA,
         PC7 = NA,
         PC8 = NA,
         PC9 = NA,
         PC10 = NA) %>% 
  select(IID, NUMATK, TOPHIGOUT, AGE1ATK, DURATION, COHORT, SEX, AGECOL, GOUT, SURICACID, ULT, PC1:PC10, PRS:PRS_std, FLAREHIGH)

ukbb_control <- ukbb_euro_pheno_prs %>% 
  filter(GOUT == FALSE,
         !is.na(PRS)) %>%
  mutate(PRS_std = (PRS - mean(PRS)) / sd(PRS),
         FLAREHIGH = NA,
         NUMATK = NA,
         TOPHIGOUT = NA,
         AGE1ATK = NA,
         DURATION = NA,
         PC1 = NA,
         PC2 = NA,
         PC3 = NA,
         PC4 = NA,
         PC5 = NA,
         PC6 = NA,
         PC7 = NA,
         PC8 = NA,
         PC9 = NA,
         PC10 = NA,
         ULT = NA) %>% 
  select(IID, NUMATK, TOPHIGOUT, AGE1ATK, DURATION, COHORT, SEX, AGECOL, GOUT, SURICACID, ULT, PC1:PC10, PRS:PRS_std, FLAREHIGH)

ukbb <- rbind(ukbb_gout, ukbb_control) %>% 
  mutate(PRS_std = (PRS - mean(PRS)) / sd(PRS))

rm(ukbb_euro_pheno_prs)


# Model results
load(here("Output/GoutvsPRS.RData"))
load(here("Output/Meta3traitsvsPRS.RData"))
load(here("Output/PooledvsPRS.RData"))
load(here("Output/CollidervsPRS.RData"))
load(here("Output/GoutvsPRS_Poly.RData"))
load(here("Output/PooledvsPRS_Poly.RData"))
load(here("Output/CollidervsPRS_Poly.RData"))
# load(here("Output/GoutvsPRS_noULT.RData"))
# load(here("Output/Meta3traitsvsPRS_noULT.RData"))
# load(here("Output/PooledvsPRS_noULT.RData"))
# load(here("Output/CollidervsPRS_noULT.RData"))
# load(here("Output/GoutvsPRS_Poly_noULT.RData"))
# load(here("Output/PooledvsPRS_Poly_noULT.RData"))
# load(here("Output/CollidervsPRS_Poly_noULT.RData"))
```

```{r functions}
# Functions 
male <- function(x){
  paste0(length(x$SEX), " (", sprintf((mean(x$SEX == "Male") * 100), fmt = "%#.1f"), ")")
}

report <- function(x) {
    if(sum(is.na(x)) != length(x)) {
      paste0(sprintf(mean(x, na.rm = TRUE), fmt = "%#.1f"), " ± ", sprintf(sd(x, na.rm = TRUE), fmt = "%#.1f"))
    } else {
      paste0("NA")
    }
}

sumreport <- function(x) {
  if(sum(is.na(x)) != length(x)){
    paste0(sum(x, na.rm = TRUE), " (", sprintf((mean(x, na.rm = TRUE) * 100), fmt = "%#.1f"), ")")
  } else {
      paste0("NA")
  }
}

transpose_df <- function(df) {
  t_df <- data.table::transpose(df)
  colnames(t_df) <- rownames(df)
  rownames(t_df) <- colnames(df)
  t_df <- t_df %>%
    rownames_to_column() %>%
    as_tibble()
  return(t_df)
}


maketable <- function(x, outname){
  assign("tmp", x %>% mutate(quintile = ntile(PRS, 5)))
  
  tmp1 <- tmp %>% filter(quintile == 1)
  tmp2 <- tmp %>% filter(quintile == 2)
  tmp3 <- tmp %>% filter(quintile == 3)
  tmp4 <- tmp %>% filter(quintile == 4)
  tmp5 <- tmp %>% filter(quintile == 5)
  
  cohortstring <- c("First PRS quintile",
                    "Second PRS quintile",
                    "Third PRS quintile",
                    "Fourth PRS quintile",
                    "Fifth PRS quintile")
  
  data_list <- list(tmp1, tmp2, tmp3, tmp4, tmp5)
  
  quintile_table <- tibble("Cohort" = cohortstring,
                           "N (% male)" = unlist(lapply(data_list, male)),
                           "Mean age (sd) (years)" = unlist(lapply(data_list, function(x) report(x$AGECOL))),
                           "Mean SU (sd) (mg/dL)" = unlist(lapply(data_list, function(x) report(x$SURICACID))),
                           "Urate-lowering therapy (%)" = unlist(lapply(data_list, function(x) sumreport(x$ULT))),
                           "Mean age at onset (sd)" = unlist(lapply(data_list, function(x) report(x$AGE1ATK))),
                           "Tophi (%)" = unlist(lapply(data_list, function(x) sumreport(x$TOPHIGOUT))),
                           "Flares > 1 (%)" = unlist(lapply(data_list, function(x) sumreport(x$FLAREHIGH))),
                           "Mean PRS (sd)" = unlist(lapply(data_list, function(x) report(x$PRS))),
                           "Mean disease duration (sd)" = unlist(lapply(data_list, function(x) report(x$DURATION))))
  
  quintile_table <- transpose_df(quintile_table) %>% 
    row_to_names(row_number = 1) %>% 
    column_to_rownames(var = "Cohort")
  
  assign(paste0(outname, "_quintiletable"), quintile_table, envir = .GlobalEnv)
}
```

```{r Table1}
cohortstring <- c("UK Biobank  - gout",
                  "UK Biobank  - controls",
                  "Aus/NZ European - gout",
                  "Aus/NZ European - controls",
                  "GlobalGout",
                  "Ardea - CLEAR/LASSO (ULT)",
                  "Ardea - CRYSTAL",
                  "Ardea - LASSO (no ULT)",
                  "Ardea - LIGHT",
                  "Pooled gout (excluding UK Biobank)")

data_list <- list(ukbb_gout, ukbb_control, anz_gout, anz_control, eurogout, ardea_clear_lasso, ardea_crystal, ardea_lasso_other, ardea_light, allgout)

table1 <- tibble("Cohort" = cohortstring, 
                 "N (% male)" = unlist(lapply(data_list, male)),
                 "Mean age (sd) (years)" = unlist(lapply(data_list, function(x) report(x$AGECOL))),
                 "Mean SU (sd) (mg/dL)" = unlist(lapply(data_list, function(x) report(x$SURICACID))),
                 "Urate-lowering therapy (%)" = unlist(lapply(data_list, function(x) sumreport(x$ULT))),
                 "Mean age at onset (sd)" = unlist(lapply(data_list, function(x) report(x$AGE1ATK))),
                 "Tophi (%)" = unlist(lapply(data_list, function(x) sumreport(x$TOPHIGOUT))),
                 "Flares > 1 (%)" = unlist(lapply(data_list, function(x) sumreport(x$FLAREHIGH))),
                 "Mean PRS (sd)" = unlist(lapply(data_list, function(x) report(x$PRS))),
                 "Mean disease duration (sd)" = unlist(lapply(data_list, function(x) report(x$DURATION))))

table1 <- transpose_df(table1) %>% 
  row_to_names(row_number = 1) %>% 
  column_to_rownames(var = "Cohort")

kable_styling(kable(table1))
```

The main cohort differences are the following:

1. Control cohorts have far fewer males, are younger, have lower SU, have lower PRS

2. UK Biobank, Aus/NZ, and GlobalGout seem to be the most similar gout cohorts

    - Aus/NZ has 7% fewer females, is ~ 4 years older, has slightly lower SU
    
    - GlobalGout has highest SU and most on ULT
    
    - All three have similar PRS
    
    - Compared to GlobalGout, Aus/NZ has 2 years older age at onset and 10% fewer active flares, but also 2 years longer disease duration and 4% more tophi
    
3. Collectively, the Ardea cohorts have fewer females, younger age at collection, higher SU, earlier age at onset, more flares, and higher PRS than GlobalGout/AusNZ

    - CRYSTAL has fewest females, earliest age at onset, all had tophi, longest disease duration
    
    - LIGHT has most females, oldest age, latest onset, highest PRS (weird), second highest SU, second most tophi, second longest disease duration

```{r PRSQuintiles_Pooled}
maketable(allgout, "AllGout")

kable_styling(kable(AllGout_quintiletable))
```

The above table shows the metrics for each quintile of PRS within the pooled gout cohort. 

1. Increasing PRS quintile associates with: more males, younger cases, increased SU, increased ULT usage, younger age at onset, slightly more tophi, more flares, longer disease duration

```{r pooledmalesvfemales}
maketable(filter(allgout, SEX == "Male"), "AllGoutMale")

kable_styling(kable(AllGoutMale_quintiletable))

maketable(filter(allgout, SEX == "Female"), "AllGoutFemale")

kable_styling(kable(AllGoutFemale_quintiletable))
```

The results of the male vs female analysis suggests the following:

1. All of the above observations are more pronounced in males vs females

2. Females seem to have the opposite disease duration pattern to males (lower PRS = longer duration)

3. The sigmoidal relationship between PRS and age at onset (also SU) seems to be more pronounced when the cohort is split into males and females

```{r individualcohorts}
# Aus/NZ cohort
maketable(anz_gout, "Aus_NZ")

kable_styling(kable(Aus_NZ_quintiletable))

# GlobalGout cohort
maketable(eurogout, "GlobalGout")

kable_styling(kable(GlobalGout_quintiletable))

# Ardea - CLEAR/LASSO cohort
maketable(ardea_clear_lasso, "Ardea_CLEAR_LASSO")

kable_styling(kable(Ardea_CLEAR_LASSO_quintiletable))

# Ardea - CRYSTAL cohort
maketable(ardea_crystal, "Ardea_CRYSTAL")

kable_styling(kable(Ardea_CRYSTAL_quintiletable))

# Ardea - LASSO (other) cohort
maketable(ardea_lasso_other, "Ardea_LASSO_other")

kable_styling(kable(Ardea_LASSO_other_quintiletable))

# Ardea - LIGHT cohort
maketable(ardea_light, "Ardea_LIGHT")

kable_styling(kable(Ardea_LIGHT_quintiletable))
```

This individual cohort analysis suggests the following:

1. In the Aus/NZ cohort, only age at onset seems to consistently decrease with PRS, though the correlations seen in the pooled cohort are present to a less consistent degree

2. In the GlobalGout cohort, the age at onset phenotype seems to be the only 
